### 2018-06-20 Writing/response_to_reviewers.tex

Besides, as the reviewer also correctly points out, we simulated our data with no missing data so this does not affect our results.
The precision in the manuscript was just describing the properties of the new CD metric.

We've clarified how missing data can be potentially treated in the methods section:
\textit{For example, for two characters X and Y for four taxa, if character \texttt{X = \{0,0,?,1\}} for the first, second, third and fourth taxa respectively and character \texttt{Y = \{0,0,1,1\}}, only the first, second and fourth taxa are compared for this character.
Note that in these simulations, no missing data was included so the characters are compared for equal number of taxa (see below).} lines @@@

\item{\textcolor{blue}{I also found the explanations confusing for what proportion of a matrix was repeat sampled. For example, in a matrix that you call maximised, n characters are replaced. But how many is that really? What is the range? I had a hard time putting a context to this and deciding how well I feel these simulations approximate reality.  Without knowing this, I was having difficulty understanding why the statement "when character correlation was high (low character differences), the topology was always the furthest away from the``normal'' topology." 
should be true - does this simply mean that the same signal is being repeated over and over again, but that signal is low information relative to the topology? To my mind, highly correlated datasets would yield congruent trees, but that these trees may be discordant with the true tree, if the resampled characters were not representative of the true signal, or highly concordant with the true tree if they were. Perhaps this not being observed is a result of not performing comparisons to the true tree?}}

We've added the proportion of replaced characters in each matrices for each ``minimised'', ``maximised'' and ``randomised'' scenarios in the supplementary material 3, figures 1, 2 and 3.
It is worth noting that there is a significant difference in terms of number of replaced characters in each scenarios between the ``maximised'' and the ``minimised'' and ``randomised'' matrices but non in terms of number of taxa or characters.

This could potentially explain some of the differences between the scenarios (i.e. why it seems better to maximise character difference rather than minimising it) but does not explain the differences in smaller matrices nor between the ``minimised'' and ``randomised'' scenarios and neither between the two inference methods (since they both use the same matrices).
Furthermore, the maximising/minimising character differences was done \textit{a priori} without any targeted number of characters to replace to make the analysis less idiosynchratic.
In other words, we decided to modify the randomly simulated matrices to globally increase/decrease character differences which is expected to decrease anyway in bigger matrices.
When using a small number of character states (i.e. two or three), the matrices with few taxa and many characters are expected to have an overall low character difference because of the smaller combinatoric possibilities for each characters.

We've added mentions to the proportion of replaced characters analysis in the supplementary in the main text at lines @@@ and added a paragraph explaining this caveat in the discussion:

%TG: TODO: added pragraph!
\textit{.} lines @@@

@@

\textit{Each replaced characters (i.e. the characters with a weight of +$1$) where randomly selected from the remaining characters using a uniform distribution (0,1).} lines @@@

\item{\textcolor{blue}{It is not obvious why generated character sequences are representative of biological sequences or sample the space overall. Were the parameters chosen for the HKY model to fit a particular evolutionary scenario? Are they meant to capture all of the space?}}

There is actually a vigorous debate surrounding how to generate biologically meaningful morphological characters \citep[][;confidential reviews and personal communications witht he authors]{OReilly20160081,goloboff2017weighted,puttick2017uncertain,GoloboffEmpirical,OReillyEmpirical}.
This ongoing debate, however, focuses mainly on inference methods where data simulation is a tangential aspect of it.
Therefore, we decided to base our simulation protocol on the empirical observations showing that half of 206 analysed datasets supported a M$k$-like model \citep[][Fig. 3]{Wright01072016} and the other half being a combination of models not explicitly described here but that are believed to be covered with the HKY-binarised model \citep{puttick2017uncertain,OReillyEmpirical}.
In our simulations, the idea was to cover a broad range of morphological characters (with variable state numbers, evolution and transition rates) in order to capture a realistic image of the ``discrete morphological character space''.

%TG: add some text to the manuscript paraphrasing this? It's kind of already written in there...

@@

\subsection{Major suggestions}

\begin{enumerate}

\item{\textcolor{blue}{I feel that the manuscript is certainly worth publishing in Systematic Biology, although I worry that the simulation procedure used by the authors limits their ability to thoroughly test of the effect of between character correlation on the accuracy of estimating phylogenetic trees. The authors note this themselves in the discussion, but I feel more should be made of this point. If the simulations utilised an explicit distribution of character covariance then I think that interpretation of the results would have been more straightforward.}}

% As noted in the introduction, the aim of this manuscript was mainly to see how do software handle correlation between discrete morphological characters.
% In other words, how can we distinguish between the complex sources of correlation in morphology

### 2018-06-28 Writing/response_to_reviewers.tex

\subsection{Major suggestions}

\begin{enumerate}

\item{\textcolor{blue}{I feel that the manuscript is certainly worth publishing in Systematic Biology, although I worry that the simulation procedure used by the authors limits their ability to thoroughly test of the effect of between character correlation on the accuracy of estimating phylogenetic trees. The authors note this themselves in the discussion, but I feel more should be made of this point. If the simulations utilised an explicit distribution of character covariance then I think that interpretation of the results would have been more straightforward.}}

As noted in the introduction, the aim of this manuscript was mainly to see how do software handle correlation between discrete morphological characters.
In other words, how can we distinguish between the complex sources of correlation in phenotype using only discrete morphological characters?
We do agree with the reviewer that a thorough analysis of the variance/covariance structure of simulated ``discretised'' characters (see reviewer 3, major suggestion \ref{hiden}) would have broad more insight into the nature of character correlation on phylogenetics inference in general (regardless of the type of character and correlation) but we feel that this is out of the scope of this study.

We have however made clearer in the introduction what were the objectives of this current manuscript:

### 2018-06-28 Writing/draft_highlighted.tex
Response updates

In the example in Table \ref{Tab:example_matrix}, the characters \textbf{C1} and \textbf{C2} are the most likely to be truly independent; characters \textbf{C3} and \textbf{C4} suffer from a a coding induced dependency; characters \textbf{C1} and \textbf{C3}/\textbf{C4} have an evolutionary induce dependency and again, characters \textbf{C2} and \textbf{C3}/\textbf{C4} are likely to be independent.
Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the character states tokens will be used in order to infer the tree.
\hl{In fact, for practicality reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} more than the other characters since this would be \textit{a priori} the characters that are the most likely to reflect the group's evolutionary history in terms of split by separating the artiodactyles from the cetaceans.
In other words, this is the correlation between characters that depicts the evolutionary history of this group.
However, the other characters depict the same history (i.e. the same split) but for the wrong reasons (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleens, the same split would still be estimated by the software).
One solution could be to weight characters \textit{a priori} but that is a bad practice since it can easily lead to the resulting tree reflecting the researcher's \textit{priori} rather than the true evolutionary history.}

This illustrates how some characters will therefore be expected to covary in non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
Yet the question has never been explored through a thorough simulation framework (although it has been tackled empirically for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence}).

How \hl{do} these correlation really affect topology?
We expect matrices with a high level of correlation to recover precise but inaccurate topologies but will matrices with low level of correlation (i.e. with high levels of homoplasy) actual cancel out the effects of correlation?
Here we formally assess the effect of discrete character's correlation using simulated data.
We propose a new distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase/decrease the overall differences or similarities between characters.
We found that overall, there is a detectable effect of character correlation on topology where an increase in character dependence results in a decrease in the ability to recover the correct topology.
These results, however, vary greatly in magnitude depending on the size of matrix and the inference method used.

>>>>

In the example in Table \ref{Tab:example_matrix}, the characters \textbf{C1} and \textbf{C2} are the most likely to be truly independent; characters \textbf{C3} and \textbf{C4} suffer from a a coding induced dependency; characters \textbf{C1} and \textbf{C3}/\textbf{C4} have an evolutionary induce dependency and again, characters \textbf{C2} and \textbf{C3}/\textbf{C4} are likely to be independent.
Yet a phylogenetic software will treat all these four characters in exactly the same way: only the sheer difference between the character states tokens will be used in order to infer the tree.
\hl{In fact, for practicality reasons, phylogenetic software do not (or rarely) distinguish between the different sources of correlation.
In this specific example, we would intuitively want to favour character \textbf{C1} and \textbf{C2} more than the other characters since this would be \textit{a priori} the characters that are the most likely to reflect the group's evolutionary history in terms of split by separating the artiodactyles from the cetaceans.
In other words, this is the correlation between characters that depicts the evolutionary history of this group.
However, the other characters depict the same history (i.e. the same split) but for the wrong reasons (note that even if \textbf{C2} was coded presence (1) or absence (0) of baleens, the same split would still be estimated by the software).
One solution could be to weight characters \textit{a priori} but that is a bad practice since it can easily lead to the resulting tree reflecting the researcher's \textit{priori} rather than the true evolutionary history.}
% \hl{Furthermore, phenotypic variations can be continuous rather than discrete and ``discretising'' such characters can have some influence on tree inference} \citep{parins2017use}.
% \hl{In this case, ``discretised'' }.
This illustrates how some characters will therefore be expected to covary in non-phylogenetic way, and that this phenomenon can reasonably be expected to mislead phylogenetic analysis.
Yet the question has never been explored through a thorough simulation framework (although it has been tackled empirically for morphological data \citealt{Davalos01072014} or molecular data \citealt{ZouConvergence}).

How \hl{do} these \hl{specific} correlation really affect topology?
% In the specific case of discrete morphological characters, how does the reduc phenotypic information and its different sources of correlation is 
% More specifically, how can phylogenetic inference software distinguish between the complex sources of correlation tha
% In other words, how can we distinguish between the complex sources of correlation in phenotype using only discrete morphological characters?
% \cite{parins2017use}
Here we formally assess the effect of discrete character's correlation using simulated data.
We propose a new distance metric to measure the difference between characters (as a proxy for these three sources of correlation as interpreted by the software) and a protocol to modify discrete morphological matrices to increase/decrease the overall differences or similarities between characters.
We found that overall, there is a detectable effect of character correlation on topology where an increase in character dependence results in a decrease in the ability to recover the correct topology.
These results, however, vary greatly in magnitude depending on the size of matrix and the inference method used.
